---
title: "LASI Social engagement joint model analysis"
author: "Peiyao Zhu"
date: "`r Sys.Date()`"
output:
  html_document:
    code_folding: hide
    highlight: pygments
    keep_md: no
    number_sections: yes
    smart: yes
    theme: sandstone
    toc: yes
    toc_float: yes
  word_document:
    toc: yes
---


```{r, include=FALSE}

library(devtools)
library(pak)
library(sp)
library(INLA)
library(INLAjoint)
library(JMbayes)
library(MASS)
library(nlme)
library(splines)
library(survival)
library(haven)
library(tidyverse)
library(dplyr)
library(car)
library(lme4)
library(sas7bdat)
library(ggplot2)
library(gtsummary)
library(ggeffects)
library(viridisLite)
library(RColorBrewer)

```

# Import data - score

This dataset contains 3,869 observations.

```{r}

data_score1 <- read_dta("/Users/zpy/University of Michigan Dropbox/Peiyao Zhu/LASI_DAD_YY/Aim 3/My data/Update 062325/transformed_102425_score.dta")

data_score1$wave <- as.numeric(data_score1$wave)
data_score1 <- data_score1 %>% mutate(wave0 = case_when(
    wave == 1 ~ 0,
    wave == 2 ~ ttey)) 

summary(data_score1$wave0)
summary(data_score1$ttey)

# delete wave 2 NA / update ttey and r1died
data_score2 <- data_score1 %>% filter(!is.na(sgcp)) %>%
  mutate(
    ttey = if_else(wave != 1, NA_real_, ttey),
    r1died = if_else(wave != 1, NA_real_, r1died)
  )%>%
  mutate(
    education = case_when(
      educ_self == 0 ~ "Never attend school",
      educ_self == 1 ~ "Have Education"
  )
)


```





# format dataset - data / variable structure

```{r}

# wide format data
data_wide_score <- read_dta("/Users/zpy/University of Michigan Dropbox/Peiyao Zhu/LASI_DAD_YY/Aim 3/My data/Update 062325/LASI_ex_102425_scoreupdate.dta")

data_wide_score <- data_wide_score %>% mutate(wave=1, wave0=0) %>%
  mutate(
    education = case_when(
      educ_self == 0 ~ "Never attend school",
      educ_self == 1 ~ "Have Education"
    ),
    status = case_when(
      is.na(sgcp2) ~ "Lost follow-up",
      !is.na(sgcp2) ~ "Have follow-up",
      TRUE ~ NA_character_
    )
  )


# Full JM Model dataset-wide
data_full_wide_score <- data_wide_score %>% 
  dplyr::select(score, score_binary, age_dad1, agesqr_dad1, male, cmastat1, creligion1, caste1, hhpcconsum, 
                cchildhl, chfiance, rural1, ceduc_l, educ_self, educ_moth, educ_fath, 
                depress1, adltot1, chronic1, wave0, prim_key, ttey, r1died, education, status)
summary(data_full_wide_score)


# Full JM Model dataset-long
data_full_score <- data_score2 %>% 
  dplyr::select(score, score_binary, age_dad1, agesqr_dad1, male, cmastat1, creligion1, caste1, hhpcconsum, 
                cchildhl, chfiance, rural1, ceduc_l, educ_self, educ_moth, educ_fath, 
                depress1, adltot1, chronic1, wave0, sgcp, bexefu, bmemory, borient, blangf, bvsp, prim_key,
                ttey, r1died, education, weight1)
summary(data_full_score)

# Convert the specified variables to factors
factor_vars <- c("male", "creligion1", "caste1", "hhpcconsum", "score_binary",
                 "rural1", "cchildhl", "chfiance", "ceduc_l", "educ_self", 
                 "educ_moth", "educ_fath", "depress1", "adltot1", "chronic1", "education")

data_full_wide_score[factor_vars] <- lapply(data_full_wide_score[factor_vars], as.factor)
data_full_score[factor_vars] <- lapply(data_full_score[factor_vars], as.factor)

# center age at mean
data_full_wide_score <- data_full_wide_score %>%
  mutate(age_dad1_centered = age_dad1 - mean(age_dad1, na.rm = TRUE))%>%
  mutate(age_dad1_centered_sq = age_dad1_centered^2)
data_full_score <- data_full_score %>%
  mutate(age_dad1_centered = age_dad1 - mean(age_dad1, na.rm = TRUE))%>%
  mutate(age_dad1_centered_sq = age_dad1_centered^2)

data_full_score$prim_key <- as.factor(data_full_score$prim_key)
data_full_wide_score$prim_key <- as.factor(data_full_wide_score$prim_key)
data_full_wide_score$r1died <- as.numeric(as.character(data_full_wide_score$r1died))
data_full_wide_score$ttey <- as.numeric(data_full_wide_score$ttey)
data_full_score$wave0 <- as.numeric(data_full_score$wave0)
data_full_wide_score$r1died <- as.factor(data_full_wide_score$r1died)

# reverse social isolation index score to social engagement score
data_full_score$socengage <- 12-data_full_score$score
data_full_wide_score$socengage <- 12-data_full_wide_score$score

# Longitudinal model score_binary reverse 
data_full_score <- data_full_score %>%
  mutate(
    socengage_binary = case_when(
      socengage >= 5 ~ "High Engagement",
      socengage <= 4 ~ "Low Engagement",
      TRUE           ~ NA_character_
    ) %>% 
      factor(levels = c("High Engagement", "Low Engagement")),

    educ_self = factor(educ_self)
  ) %>%

  mutate(
    educ_self = fct_relevel(educ_self, "1")
  )

# Survival model score_binary reverse 
data_full_wide_score <- data_full_wide_score %>%
  mutate(
   socengage_binary = case_when(
      socengage >= 5 ~ "High Engagement",
      socengage <= 4 ~ "Low Engagement",
      TRUE           ~ NA_character_
    ) %>% 
      factor(levels = c("High Engagement", "Low Engagement")),
    
    educ_self = factor(educ_self)
  ) %>%
  mutate(
    educ_self = fct_relevel(educ_self, "1")
  )

data_full_score$ttey[data_full_score$ttey == 0] <- 0.001
data_full_wide_score$ttey[data_full_wide_score$ttey == 0] <- 0.001
data_full_score$socengage_binary <- relevel(data_full_score$socengage_binary, ref = "Low Engagement")
data_full_wide_score$socengage_binary <- relevel(data_full_wide_score$socengage_binary, ref = "Low Engagement")

write_dta(data_full_score, "Path")

write_dta(data_full_wide_score, "Path")

```



# Sample Distribution by Education
```{r}

# Creating the summary table
data_full_wide_score %>% 
  dplyr::select(socengage, socengage_binary, education, age_dad1, agesqr_dad1, age_dad1_centered, age_dad1_centered_sq, male, cmastat1,
                creligion1, caste1, hhpcconsum, cchildhl, chfiance, rural1, ceduc_l, educ_moth, educ_fath,
                depress1, adltot1, chronic1, score, score_binary, educ_self, r1died, status) %>% 
  gtsummary::tbl_summary(
    by = education, 
    type = score ~ "continuous2",
    statistic = list(
      all_continuous() ~ "{mean} ({sd})",
      all_categorical() ~ "{n} ({p}%)"
    ),
    digits = all_continuous() ~ 2,
    label = list(
      score ~ "Social Isolation Index Score",
      score_binary ~ "Social Isolation Index Binary",
      socengage ~ "Social Engagement Score",
      socengage_binary ~ "Social Engagement Index Binary",
      age_dad1 ~ "Age",
      agesqr_dad1 ~ "Age Squared",
      age_dad1_centered ~ "Age (Mean)",
      age_dad1_centered_sq ~ "Age (Mean) Squared",
      male ~ "Sex",
      cmastat1 ~ "Marital Status",
      creligion1 ~ "Religion",
      caste1 ~ "Caste",
      hhpcconsum ~ "Household Per Capita Consumption",
      cchildhl ~ "Childhood Health",
      chfiance ~ "Childhood Finance",
      rural1 ~ "Rural Area",
      ceduc_l ~ "Education Level",
      educ_moth ~ "Mother's Education",
      educ_fath ~ "Father's Education",
      depress1 ~ "Depression",
      adltot1 ~ "ADL",
      chronic1 ~ "Chronic Conditions",
      educ_self ~ "Education Level",
      r1died ~ "Survival"
    )
  ) %>%  
  add_overall() %>% 
  add_p(list(all_continuous() ~ "t.test"),
        pvalue_fun = ~style_pvalue(.x, digits = 3))



data_full_wide_score %>% 
  dplyr::select(socengage, socengage_binary, education, age_dad1, agesqr_dad1, age_dad1_centered, age_dad1_centered_sq, male, cmastat1,
                creligion1, caste1, hhpcconsum, cchildhl, chfiance, rural1, ceduc_l, educ_moth, educ_fath,
                depress1, adltot1, chronic1, score, score_binary, educ_self, r1died, status) %>% 
  gtsummary::tbl_summary(
    by = socengage_binary, 
    type = score ~ "continuous2",
    statistic = list(
      all_continuous() ~ "{mean} ({sd})",
      all_categorical() ~ "{n} ({p}%)"
    ),
    digits = all_continuous() ~ 2,
    label = list(
      education ~ "Education Level",
      age_dad1 ~ "Age",
      agesqr_dad1 ~ "Age Squared",
      age_dad1_centered ~ "Age (Mean)",
      age_dad1_centered_sq ~ "Age (Mean) Squared",
      male ~ "Sex",
      cmastat1 ~ "Marital Status",
      creligion1 ~ "Religion",
      caste1 ~ "Caste",
      hhpcconsum ~ "Household Per Capita Consumption",
      cchildhl ~ "Childhood Health",
      chfiance ~ "Childhood Finance",
      rural1 ~ "Rural Area",
      ceduc_l ~ "Education Level",
      educ_moth ~ "Mother's Education",
      educ_fath ~ "Father's Education",
      depress1 ~ "Depression",
      adltot1 ~ "ADL",
      chronic1 ~ "Chronic Conditions",
      educ_self ~ "Education Level",
      r1died ~ "Survival"
    )
  ) %>%  
  add_overall() %>% 
  add_p(list(all_continuous() ~ "t.test"),
        pvalue_fun = ~style_pvalue(.x, digits = 3))


```


# Joint model - INLAjoint package


```{r}

# ID continuous

data_full_score <- as.data.frame(data_full_score)
data_full_wide_score <- as.data.frame(data_full_wide_score)

id_map <- data.frame(
  prim_key = unique(as.character(data_full_score$prim_key)),
  id_jm = seq_along(unique(as.character(data_full_score$prim_key)))
)

data_full_score$id_jm <- id_map$id_jm[
  match(as.character(data_full_score$prim_key), id_map$prim_key)
]

data_full_wide_score$id_jm <- id_map$id_jm[
  match(as.character(data_full_wide_score$prim_key), id_map$prim_key)
]


```


# General cognition - social engagement score - binary
```{r}

data_full_wide_score <- data_full_wide_score %>%
  mutate(
    r1died = as.integer(as.character(r1died))
  )

data_full_score <- data_full_score %>%
  mutate(
    r1died = as.integer(as.character(r1died))
  )

## Longitudinal model formula (lform)
### LME general domain - model interaction: age, religion, childhood health, rural residence
lform_sgcp_score_model_bi <- sgcp ~ socengage_binary + wave0 + wave0 * socengage_binary + age_dad1_centered + age_dad1_centered_sq + male + creligion1 + caste1 + hhpcconsum + cchildhl + 
                         chfiance + rural1 + educ_self + educ_moth + educ_fath + depress1 + 
                         adltot1 + chronic1 + educ_self*socengage_binary + wave0*educ_self + wave0:socengage_binary*educ_self +
                         wave0*age_dad1_centered + wave0*creligion1 + wave0*cchildhl + 
                         wave0*rural1 + (1 + wave0 | id_jm)

## Survival model formula (sform)
sform_sgcp_score_model_bi <- inla.surv(ttey, r1died) ~ socengage_binary + age_dad1_centered + age_dad1_centered_sq + male + 
                         creligion1 + caste1 + hhpcconsum + cchildhl + chfiance + rural1 + educ_self + educ_moth + 
                         educ_fath + depress1 + adltot1 + chronic1 + educ_self*socengage_binary

## Fit the joint model with INLAjoint
joint_sgcp_score_model_bi <- joint(
  formSurv = sform_sgcp_score_model_bi,
  formLong = lform_sgcp_score_model_bi, 
  family = "normal",
  dataLong = data_full_score,
  dataSurv = data_full_wide_score,
  id = "id_jm",
  timeVar = "wave0",
  assoc = "SRE",
  basRisk = "rw2",
  NbasRisk = 10,
  control = list(int.strategy = "eb")
)

# Summarize models
summary(joint_sgcp_score_model_bi,hr=TRUE)


```


# executive function - social engagement score - binary
```{r}

### LME Executive fuction domain - model interaction: caste, religion, childhood health, rural residence
lform_exe_score_model_bi <- bexefu ~ socengage_binary + wave0 + wave0*socengage_binary + age_dad1_centered + age_dad1_centered_sq + male + 
                         creligion1 + caste1 + hhpcconsum + cchildhl + chfiance + rural1 + educ_self + educ_moth + 
                         educ_fath + depress1 + adltot1 + chronic1 + educ_self*socengage_binary + wave0*educ_self + 
                         wave0:socengage_binary*educ_self + wave0*caste1 + wave0*creligion1 + wave0*cchildhl + 
                         wave0*rural1 + (1 + wave0 | id_jm)

## Survival model formula (sform)
sform_exe_score_model_bi <- inla.surv(ttey, r1died) ~ socengage_binary + age_dad1_centered + age_dad1_centered_sq + male + 
                         creligion1 + caste1 + hhpcconsum + cchildhl + chfiance + rural1 + educ_self + educ_moth + 
                         educ_fath + depress1 + adltot1 + chronic1 + educ_self*socengage_binary

## Fit the joint model with INLAjoint
joint_exe_score_model_bi <- joint(
  formSurv = sform_exe_score_model_bi,
  formLong = lform_exe_score_model_bi, 
  family = "normal",
  dataLong = data_full_score,
  dataSurv = data_full_wide_score,
  id = "id_jm",
  timeVar = "wave0",
  assoc = "SRE",
  basRisk = "rw2",
  NbasRisk = 10,
  control = list(int.strategy = "eb")
)

# Summarize models
summary(joint_exe_score_model_bi,hr=TRUE)

```




# memory function - social engagement score -  binary
```{r}
summary(data_full_wide_score$socengage_binary)
## Longitudinal model formula (lform)
### LME memory domain - model interaction: age, religion, childhood health, ADL, rural residence, chronic conditions
lform_mem_score_model_bi <- bmemory ~ socengage_binary + wave0 + wave0 * socengage_binary + age_dad1_centered + age_dad1_centered_sq + male + 
                         creligion1 + caste1 + hhpcconsum + cchildhl + 
                         chfiance + rural1 + educ_self + educ_moth + educ_fath + depress1 + 
                         adltot1 + chronic1 + educ_self*socengage_binary + wave0*educ_self + wave0:socengage_binary*educ_self +
                         wave0*age_dad1_centered + wave0*adltot1 + wave0*creligion1 + wave0*cchildhl + 
                         wave0*rural1 + wave0*chronic1 + (1 + wave0 | id_jm)

## Survival model formula (sform)
sform_mem_score_model_bi <- inla.surv(ttey, r1died) ~ socengage_binary + age_dad1_centered + age_dad1_centered_sq + male + 
                         creligion1 + caste1 + hhpcconsum + cchildhl + chfiance + rural1 + educ_self + educ_moth + 
                         educ_fath + depress1 + adltot1 + chronic1 + educ_self*socengage_binary

## Fit the joint model with INLAjoint
joint_mem_score_model_bi <- joint(
  formSurv = sform_mem_score_model_bi,
  formLong = lform_mem_score_model_bi, 
  family = "normal",
  dataLong = data_full_score,
  dataSurv = data_full_wide_score,
  id = "id_jm",
  timeVar = "wave0",
  assoc = "SRE",
  basRisk = "rw2",
  NbasRisk = 10,
  control = list(int.strategy = "eb")
)

# Summarize models
summary(joint_mem_score_model_bi,hr=TRUE)

```




# language function - social engagement score - binary
```{r}

### LME language domain - model interaction: age, religion, childhood health, childhood financial status, rural residence, household consumption
lform_lan_score_model_bi <- blangf ~ socengage_binary + wave0 + wave0 * socengage_binary + age_dad1_centered + age_dad1_centered_sq + male + 
                         creligion1 + caste1 + hhpcconsum + cchildhl + 
                         chfiance + rural1 + educ_self + educ_moth + educ_fath + depress1 + 
                         adltot1 + chronic1 + educ_self*socengage_binary + wave0*educ_self + 
                         wave0:socengage_binary*educ_self +
                         wave0*age_dad1_centered + wave0*creligion1 + wave0*cchildhl + 
                         wave0*rural1 + wave0*hhpcconsum + wave0*chfiance + (1 + wave0 | id_jm)

## Survival model formula (sform)
sform_lan_score_model_bi <- inla.surv(ttey, r1died) ~ socengage_binary + age_dad1_centered + age_dad1_centered_sq + male +
                         creligion1 + caste1 + hhpcconsum + cchildhl + chfiance + rural1 + educ_self + educ_moth + 
                         educ_fath + depress1 + adltot1 + chronic1 + educ_self*socengage_binary

## Fit the joint model with INLAjoint
joint_lan_score_model_bi <- joint(
  formSurv = sform_lan_score_model_bi,
  formLong = lform_lan_score_model_bi, 
  family = "normal",
  dataLong = data_full_score,
  dataSurv = data_full_wide_score,
  id = "id_jm",
  timeVar = "wave0",
  assoc = "SRE",
  basRisk = "rw2",
  NbasRisk = 10,
  control = list(int.strategy = "eb")
)

# Summarize models
summary(joint_lan_score_model_bi,hr=TRUE)

```




# visuospatial function - social engagement score - binary
```{r}

### LME visuospatial domain - model interaction: age, gender, religion, childhood health
lform_visu_score_model_bi <- bvsp ~ socengage_binary + wave0 + wave0 * socengage_binary + age_dad1_centered + age_dad1_centered_sq + male + 
                         creligion1 + caste1 + hhpcconsum + cchildhl + 
                         chfiance + rural1 + educ_self + educ_moth + educ_fath + depress1 + 
                         adltot1 + chronic1 + educ_self*socengage_binary + wave0*educ_self + 
                         wave0*socengage_binary*educ_self +
                         wave0*age_dad1_centered + wave0*male + wave0*creligion1 + wave0*cchildhl + (1 + wave0 | id_jm)

## Survival model formula (sform)
sform_visu_score_model_bi <- inla.surv(ttey, r1died) ~ socengage_binary + age_dad1_centered + age_dad1_centered_sq + male + 
                         creligion1 + caste1 + hhpcconsum + cchildhl + chfiance + rural1 + educ_self + educ_moth + 
                         educ_fath + depress1 + adltot1 + chronic1 + educ_self*socengage_binary

## Fit the joint model with INLAjoint
joint_visu_score_model_bi <- joint(
  formSurv = sform_visu_score_model_bi,
  formLong = lform_visu_score_model_bi, 
  family = "normal",
  dataLong = data_full_score,
  dataSurv = data_full_wide_score,
  id = "id_jm",
  timeVar = "wave0",
  assoc = "SRE",
  basRisk = "rw2",
  NbasRisk = 10,
  control = list(int.strategy = "eb")
)

# Summarize models
summary(joint_visu_score_model_bi,hr=TRUE)

```



